\documentclass{llncs}

\usepackage{fancyvrb}
\usepackage{tikz}
\usepackage{boxedminipage}
\usepackage{listings}
\usepackage{lstsparql}
\usepackage{lstcsv}
\usepackage{comment}
\usepackage{xcolor}
\usepackage{hyperref}
\input{macro}

\title{The \solasote{} ontology\\
(Draft as of \today)}

\author{Ralf L\"ammel \and Martin Leinberger \and Andrei Varanovich}

\institute{Fachbereich Informatik, Universit\"at Koblenz-Landau, Germany}

\begin{document}

\maketitle

\begin{abstract}
  This document describes and exercises the \solasote{} ontology for
  software languages, software technologies, and software concepts, as
  they are relevant for programming and software engineering. Evidence
  from the software chrestomathy of the 101 project is used to
  demonstrate the ontology.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

\todo{Ralf}{Write this section.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Basics of \solasote}

\subsection{Entity types of \solasote}

Each individual of \solasote{} is of one of the following entity
types:

\completeOutputTabular{{l|l}}{\textbf{Type} & \textbf{Headline}}{individuals}

\todo{all}{The above list is still wrong.}

\sparql{individuals}

\todo{all}{We also need to support some sort of headline in the .json
  file which needs to arrive in some way in the triplestore.}

\partialOutput{7}{individuals}

Let's have a look at illustrative individuals for some of the entity
types. The following SPARQL query retrieves all software languages of
the ontology:

\sparql{languages}

\sparqlnote{\todo{Ralf}{Explain new SPARQL constructs.}}

\noindent
Here is the first few languages sorted alphabetically:

%\partialOutputTabular{{l|l}}{\textbf{Language} & \textbf{Headline}}{7}{languages}

\noindent
\solasote{} covers software technologies, too. Here, is a SPARQL query
that retrieves all technologies of the ontology:

\sparql{technologies}

\noindent
Here is the first few technologies sorted alphabetically:

\partialOutput{7}{technologies}

\noindent
\solasote{} covers software concepts, too. Here, is a SPARQL query
that retrieves all concepts of the ontology:

\sparql{concepts}

\noindent
Here is the first few concepts sorted alphabetically:

\partialOutput{7}{concepts}

\todo{Ralf}{What to do about names starting in `101'.}

\todo{Ralf}{Clearly, some of the listed names are not proper concept names.}

Likewise, other types of entities may be explored, e.g.:

\begin{itemize}
\item Features
\item Contributions
\item Contributors
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Query/output series}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Classification of \solasote's entities}

\solasote{} organizes classifiers for software languages,
technologies, and concepts. In fact, it uses a class hierarchy (in the
sense of rdfs:subClassOf) with instances at the leaves (in the sense
of rdf:type) for classification.

All classifiers are concepts, but not all concepts are
classifiers. Each classifier is associated with a concept of the same
name with a different prefix. We see this situation by looking at the
first few concept in alphabetical order as well as the first few
classifiers, again in alphabetical order and with associated concepts: 

\sparql{concepts}

\partialOutput{7}{concepts}

\sparql{classifiers}

\partialOutput{7}{classifiers}

\noindent
Let us see now how classifiers are instantiated. Here is how we ask
whether \uri{lang:Java} is a (software) language:

\sparql{javaOfTypeLanguage}

\noindent
Evaluation of this query literally returns `true':

\completeOutput{javaOfTypeLanguage}

\noindent
Instead, we may also ask for all types of \uri{lang:Java}:

\sparql{typesOfJava}

\noindent
\uri{lang:Java} has indeed a number of types:

\todo{Ralf}{Why is Java of type Resource?}

\completeOutput{typesOfJava}

\noindent
These types can be explained in the following way:
%
\begin{description}
%
\item[\uri{onto:Language}] This type stems from the entity's namespace
  on the \ooo{wiki}. That is, `Java' belongs to \ooo{wiki}'s namespace
  `Language' and thus, it gets associated with the type \uri{onto:Language}
  by the mapping from \ooo{wiki} to \solasote
%
\item[\uri{onto:OO\_programming\_language}] This type stems from a
  semantic property stated on \ooo{wiki}'s page for \uri{lang:Java}. That
  is, the
  page\footnote{\url{http://101companies.org/wiki/Language:Java}}
  contains markup for a relevant semantic property as follows:
\begin{center}
\begin{BVerbatim}
[[instanceOf::OO programming language]]
\end{BVerbatim}
\end{center}
\noindent
In this manner, the wiki directly declares `Java' to be of type `OO
programming language'.
%
\item[\uri{onto:Instrument}, \uri{onto:Entity}] These are supertypes
  of \uri{onto:Language}, as declared explicitly by the general class
  hierarchy for \solasote. Thus, these types are listed because they
  are inferred by the \emph{rdfs:subClassOf} hierarchy of \solasote.
%
\item[\uri{onto:Language}, \uri{onto:Programming\_language}]
  These are supertypes of \uri{onto:OO\_programming\_language}, as
  declared by semantic `isA' properties on the \ooo{wiki}. For
  instance, the
  page for \uri{onto:OO\_programming\_language}\footnote{\url{http://101companies.org/wiki/OO_programming_language}}
  contains markup for a relevant semantic property as follows:
\begin{center}
\begin{BVerbatim}
[[isA::Programming language]]
\end{BVerbatim}
\end{center}
\noindent
Thus, these types are listed because they are inferred by the
\emph{rdfs:subClassOf} hierarchy of \solasote. The type
\uri{onto:Language} is a shorthand for what is called `software
language' on the wiki. The type \uri{onto:Programming\_language} is a
proper classifier of (software) languages (just as much as
\uri{onto:OO\_programming\_language}).
%
\item[\uri{onto:WikiPage}] This type conveys that the resource at hand
  corresponds to a wiki page on the \ooo{wiki}. This is basically
  true for all resources of type \uri{onto:Entity}. The type factors
  all more representational properties of entities; see
  \S\ref{S:mapping}.
%
%\item[\uri{rdfs:Resource}] This is the root type of all resources
%  based on RDFS' underlying type system.
\end{description}
%
We can also query the supertypes of a given type. For instance, let's
query the supertypes of \uri{onto:OO\_programming\_language}, as we
have encountered this type previously:

\todo{Martin}{Once the name mapping of 26 Aug is in place, then
  uri{rdfs:Resource} should be gone from the output; ditto \uri{onto:Software\_language}.}

\sparql{supertypesOfOoProgrammingLanguage}

\noindent
These are all the supertypes of \uri{onto:OO\_programming\_language};
please note that all the types but \uri{onto:Programming\_language}
are inferred and that inference clearly assumes reflexivity, since
\uri{onto:OO\_programming\_language} is returned as a supertype of itself:

\completeOutput{supertypesOfOoProgrammingLanguage}

\noindent
Alternatively, we can also query for all the instances of a certain
class. For instance, we can ask for all the OO programming languages:

\sparql{instancesOfOoProgrammingLanguage}

\noindent
These are some of the OO programming languages:

\partialOutput{7}{instancesOfOoProgrammingLanguage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Mapping from \ooo{wiki} to \solasote}
\label{S:mapping}

\todo{Ralf}{Write this section.}

\sparql{haskellHeadline}

\sparql{haskellWikiLink}

\completeOutput{haskellHeadline}

\completeOutput{haskellWikiLink}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The schema of \solasote}

\todo{Ralf}{Write this section.}

\sparql{properties}

\todo{Martin and Andrei}{Integrate 101's properties into
  triplestore. This requires a mapping from the DSL to RDFS/Owl.}

\completeOutput{properties}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Popularity of \solasote's entities}

We may be interested in the popularity of different kinds of entities
in the ontology. For instance, we may use the number of distinct
contributions referring to an entity for ranking all entities of a
certain kind (such as languages, technologies, or concepts). 

\sparql{popularLanguages}

\sparqlnote{\todo{Ralf}{Explain new SPARQL constructs.}}

\partialOutput{7}{popularLanguages}

\sparql{popularTechnologies}

\partialOutput{7}{popularTechnologies}

\todo{Ralf}{Why is contrib:haskellEngineer a popular technology?}

\sparql{popularConcepts}

\partialOutput{7}{popularConcepts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Validation of \solasote}

\todo{Ralf}{Write this section.}

\todo{Martin}{Complete sandbox upload on 101worker.}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Concluding remarks}

\todo{Ralf}{Write this section.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{comment}

\input{associationOfDevelopersPages}

\input{baselineContribution}

\input{conceptsToParadigm}

\input{developerToParadigm}

\input{disjointness}

\input{knowledgeIslands}

\input{languageContributionNoOfFeatures}

\input{languageFeatureContributionNoOfFeatures}

\input{languageFeatureContributionSumOfLangsAndTechs}

\input{progressionOfLectureTopics}

\input{technologyContributionNoOfFeatures}

\input{togetherness}

\end{comment}

\end{document}